---
title: "AUC Examples"
author: "David Elsheimer"
date: "2/26/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Monte Carlo 

The following is sample code showing estimation of area under the curve for a standard normal distribution.

Note the true value of $AUC(-1,1)= \int_a^b \frac{1}{\sqrt{2\pi}}e^{-\frac{x^2}{2}}dx = \Phi(1) - \Phi(-1)$. This can easily be calculated in R using $\texttt{pnorm}$. The true value is found to be 0.6826895.

A vector of values of the estimated AUC for $n \in \{10,100,1000,10000\}$ is provided below.

```{r}
library(ggplot2)
set.seed(1002)
aucest<- c(RStudio2020::estimate_auc(10,-1,1),
RStudio2020::estimate_auc(1e2,-1,1),
RStudio2020::estimate_auc(1e3,-1,1),
RStudio2020::estimate_auc(1e4,-1,1))
num<- c(10,100,1000,10000)
plotdata<- as.data.frame(cbind(num,aucest))
plotdata
ggplot(data = plotdata, aes(x=num, y=aucest)) + geom_point()+
    geom_line(linetype="dashed")+geom_hline(yintercept=0.6826895, color = "red")
```

 Let $Z_1, Z_2, ..., Z_n$ denote $n$ i.i.d random variables such that $\lvert Z_i \rvert \leq 1$ for $i = 1, ..., n.$ Let $\overline{Z}$ denote the sample mean of $Z_i$ and $\mu$ denote the mean of $Z_1$. Recall that the Hoeffding bound tell us that

$$
P(\lvert \overline{Z} - \mu \rvert \geq t) \leq 2 \exp(-2 n t^2).
$$

Here, $Z_i = I(Y_i \leq f(X_i))$. The $Z_i$'s are i.i.d., and $|Z_i| \leq 1$. Thus a Hoeffding bound is applicable here. $\bar{Z} = \frac{1}{n}\sum_{i=1}^nZ_i$. Here $E(Z_i) = P(Y_i \leq f(X_i))$. Due to $Y_i$'s being i.i.d, as well as $X_i$'s being i.i.d., this can be further generalized as $P(Y\leq f(X))$.

Further note that $P(Y\leq f(X)) = \frac{S}{c(b-a)}$, where $a,b$ are the bounds for $X_i$, and $c$ is the maximum density of $f(X)$, and $S$ is the area under the curve.

Therefore, $S = c(b-a)P(Y \leq f(X))\approx c(b-a)\bar{Z}$. 

Applying the Hoeffding bound, $P(|c(b-a)\bar{Z} - c(b-a)P(Y\leq f(X))| \geq c(b-a)t)\leq 2\exp(-2n(c(b-a)t)^2)$.

Applying what has been demonstrated above, we can assess the correctness of `estimate_auc` using the following R script.


```{r, echo=TRUE, warning=FALSE, message=FALSE}
library(tidyr)
library(dplyr)
library(forcats)
library(ggplot2)
set.seed(1002)

#Simulating estimates for different levels of n
#Creating boxplots at each level
num<- c(10,50,100,500,1000,5000,10000)
auc_eval <- replicate(sapply(num,a=-1,b=1, RStudio2020::estimate_auc), n =1000)
auc_data <- as.data.frame(t(auc_eval)) 
colnames(auc_data) <- c("10", "50", "100", "500","1000","5000","10000")
longauc <- gather(auc_data,key=num)
longauc$num<-factor(longauc$num, levels=c("10", "50", "100", "500","1000","5000","10000"))
ggplot(data = longauc, aes(x=num, y=value)) + geom_boxplot() +
    geom_hline(yintercept=0.6826895, color = "red") + 
  xlab("Number of darts thrown") + 
  ylab("Area under curve via MC")


#Checking that the probablities will be less than the bounds
probbound<-function(n,t){
  2*exp(-2*n*t**2)
}

true_y <- pnorm(1)-pnorm(-1)
longauc$count_tol1 <- 1- as.numeric(between(longauc$value,true_y-0.1,true_y+0.1))
longauc$count_tol2 <- 1-as.numeric(between(longauc$value,true_y-0.05,true_y+0.05))
longauc$count_tol3 <- 1-as.numeric(between(longauc$value,true_y-0.025,true_y+0.025))
probapprox <- longauc %>% group_by(num)%>%
  summarise(mean(count_tol1), mean(count_tol2),mean(count_tol3))
orderedprob <- probapprox
orderedprob$probtol1 <-sapply(num, t=0.1, probbound)
orderedprob$probtol2 <-sapply(num, t=0.05, probbound)
orderedprob$probtol3 <- sapply(num, t=0.025, probbound)
knitr::kable(orderedprob, col.names = c("N", "$\\hat{P}(est\\notin$ tol-int 1)",
                           " $\\hat{P}(est\\notin$ tol-int 2)", 
                           "$\\hat{P}(est\\notin$ tol-int 3)", 
                           "Tol1 True prob",
                           "Tol2 True prob",
                           "Tol3 True prob"))


```

A modified AUC function was also created which returns an estimated AUC with a specified tolerance of probability `p`.


```{r, echo=TRUE}
RStudio2020::estimate_auc_tol(-1,1,0.5,0.5)
RStudio2020::estimate_auc_tol(-1,1,1,0.1)
RStudio2020::estimate_auc_tol(-1,1,0.05,0.05)
```

As the probability decreases, the estimate requires larger and larger n, and the estimate trends towards the true AUC.


Suppose instead of the $Y$-space is changed, such that the spac eis now $[a, b]\times[0,c]$ for $c \geq 1/\sqrt{2\pi}$. With the fraction of points under the curve still being what is estimated.

The modification is simple. The change here is that $Y \sim U(0,c)$. This will change the estimator with respect to how the height of the canvas changes. Notice that  for fixed $a,b$ where $b>a$, X is bounded between $(\frac{1}{\sqrt{2\pi}}e^{-b^2/2}, \frac{1}{\sqrt{2\pi}})$. As c increases, $P(A_1) \approx \frac{1}{n}\sum_{i=1}^nI(A_i)$ will decrease because it will become increasingly likely that there will be values of $y_i>f(x_i)$. Thus the quality of the estimator will decrease for greater values of $c$. This can be coded fairly easily, as seen above and in the ackage. A condition is included to return errors for canvases shorter than the minimum height if $\frac{1}{\sqrt{2\pi}}$.

```{r, echo = TRUE, error=TRUE}


set.seed(1002)
#demonstrating the behavior
RStudio2020::estimate_auc_modified(10,-1,1, c= .1)
RStudio2020::estimate_auc_modified(10,-1,1)
RStudio2020::estimate_auc_modified(10,-1,1, c= 3)
RStudio2020::estimate_auc_modified(10,-1,1, c= 4)

```

