---
title: "Power Method Examples"
author: "David Elsheimer"
date: "2/26/2020"
header-includes:
  - \usepackage{bm}
  - \newcommand{\Real}{\mathbb{R}}
  - \newcommand{\sweep}{{\bf sweep}\,}
  - \newcommand{\dom}{{\bf dom}\,}
  - \newcommand{\sign}{{\bf sign}\,}
  - \newcommand{\Tra}{^{\sf T}} 
  - \newcommand{\Inv}{^{-1}}
  - \def\vec{\mathop{\rm vec}\nolimits}
  - \def\sweep{\mathop{\rm sweep}\nolimits}
  - \newcommand{\diag}{\mathop{\rm diag}\nolimits}
  - \newcommand{\tr}{\operatorname{tr}}
  - \newcommand{\epi}{\operatorname{epi}}
  - \newcommand{\V}[1]{{\bm{\mathbf{\MakeLowercase{#1}}}}} 
  - \newcommand{\VE}[2]{\MakeLowercase{#1}_{#2}} 
  - \newcommand{\Vn}[2]{\V{#1}^{(#2)}} 
  - \newcommand{\Vtilde}[1]{{\bm{\tilde \mathbf{\MakeLowercase{#1}}}}}
  - \newcommand{\Vhat}[1]{{\bm{\hat \mathbf{\MakeLowercase{#1}}}}} 
  - \newcommand{\VtildeE}[2]{\tilde{\MakeLowercase{#1}}_{#2}}
  - \newcommand{\M}[1]{{\bm{\mathbf{\MakeUppercase{#1}}}}}
  - \newcommand{\ME}[2]{\MakeLowercase{#1}_{#2}} 
  - \newcommand{\Mtilde}[1]{{\bm{\tilde \mathbf{\MakeUppercase{#1}}}}} 
  - \newcommand{\Mhat}[1]{{\bm{\hat \mathbf{\MakeUppercase{#1}}}}} 
  - \newcommand{\Mcheck}[1]{{\bm{\check \mathbf{\MakeUppercase{#1}}}}} 
  - \newcommand{\Mbar}[1]{{\bm{\bar \mathbf{\MakeUppercase{#1}}}}} 
  - \newcommand{\Mn}[2]{\M{#1}^{(#2)}} 
output: pdf_document
---



## Power Method Examples

Usage of the `power_method_dense` function to compute the largest singular value of the following matrices, comparing results with what is provided by the `svd` function in R. 

```{r, echo=TRUE}
set.seed(12345)
n <- 1e3
p <- 2e3
A <- matrix(rnorm(n*p), n, p)
B <- matrix(rnorm(n*p), n, p)
C <- matrix(rnorm(n*p), n, p)

x<-rnorm(p)
k<-1000
tol<-1e-6
xapproxA<- RStudio2020::power_method_dense(A,x,k,tol)
xapproxB<- RStudio2020::power_method_dense(B,x,k,tol)
xapproxC<- RStudio2020::power_method_dense(C,x,k,tol)

round(norm(A%*%xapproxA,type="2"),3)
round(max(svd(A)$d),3)

round(norm(B%*%xapproxB,type="2"),3)
round(max(svd(B)$d),3)

round(norm(C%*%xapproxC,type="2"),3)
round(max(svd(C)$d),3)
```

Usage of the `power_method_sparse` function to compute the largest singular value of the following matrices, comparing results with what is provided by the `svd` function in R. 

```{r, echo=TRUE}
library(Matrix)
set.seed(12345)
n <- 1e3
nnz <- 0.1*n
ix <- sample(1:n, size = nnz, replace = FALSE)
A <- Matrix(0, nrow=n, ncol=n, sparse=TRUE)
A[ix] <- rnorm(nnz)
A[1,1] <- 10

ix <- sample(1:n, size = nnz, replace = FALSE)
B <- Matrix(0, nrow=n, ncol=n, sparse=TRUE)
B[ix] <- rnorm(nnz)
B[1,1] <- 10

ix <- sample(1:n, size = nnz, replace = FALSE)
C <- Matrix(0, nrow=n, ncol=n, sparse=TRUE)
C[ix] <- rnorm(nnz)
C[1,1] <- 10

x<-rnorm(ncol(A))
k<-1000
tol<-1e-6

xapproxA<- RStudio2020::power_method_sparse(A,x,k,tol)
xapproxB<- RStudio2020::power_method_sparse(B,x,k,tol)
xapproxC<- RStudio2020::power_method_sparse(C,x,k,tol)

round(norm(A%*%xapproxA,type="2"),3)
round(max(svd(as.matrix(A))$d),3)

round(norm(B%*%xapproxB,type="2"),3)
round(max(svd(as.matrix(B))$d),3)

round(norm(C%*%xapproxC,type="2"),3)
round(max(svd(as.matrix(C))$d),3)
```

\newpage

Usage of the `power_method_low_rank` function to compute the largest singular value of the matrices $\M{U}_1\M{V}_1\Tra, \M{U}_2\M{V}_2\Tra,$ and $\M{U}_3\M{V}_3\Tra$ defined below, compared to what is provided by the `svd` function in R. 

```{r, echo=TRUE}
set.seed(12345)
n <- 1e3
k <- 10
U1 <- V1 <- matrix(rnorm(n*k), n, k)

U2 <- V2 <- matrix(rnorm(n*k), n, k)

U3 <- V3 <- matrix(rnorm(n*k), n, k)


x<-rnorm(n)
k<-1000
tol<-1e-6
xapproxU1V1<- RStudio2020::power_method_low_rank(U1,V1,x,k,tol)
xapproxU2V2<- RStudio2020::power_method_low_rank(U2,V2,x,k,tol)
xapproxU3V3<- RStudio2020::power_method_low_rank(U3,V3,x,k,tol)

norm((U1%*%t(V1))%*%xapproxU1V1,type="2")
max(svd(U1%*%t(V1))$d)

norm((U2%*%t(V2))%*%xapproxU2V2,type="2")
max(svd(U2%*%t(V2))$d)

norm((U3%*%t(V3))%*%xapproxU3V3,type="2")
max(svd(U3%*%t(V3))$d)
```

The largest value of the following matrix is estimated below.

$$
\M{A} = \M{S} + \V{u}\V{u}\Tra.
$$

```{r, echo=TRUE, error=TRUE,warning=TRUE}
set.seed(12345)
n <- 1e7
nnz <- 1e-5*n
ix <- sample(1:n, size = nnz, replace = FALSE)
S <- Matrix(0, nrow=n, ncol=n, sparse=TRUE)
S[ix] <- rnorm(nnz)
S[1,1] <- 10

u <- matrix(rnorm(n), ncol=1)



x<-rnorm(n)
k<-1000
tol<-1e-6

max(svd(S+u%*%t(u))$d)
RStudio2020::power_method_dense(S+u%*%t(u),x,k,tol)
xapproxsplr<-RStudio2020::power_method_sparse_plus_low_rank(S,u,u,x,k,tol)

###Largest value
sqrt(t(xapproxsplr)%*%as.matrix((Matrix::t(S)%*%S)%*%xapproxsplr +
                     u%*%(t(u)%*%u)%*%(t(u)%*%xapproxsplr) +
                     u%*%(t(u)%*%(S%*%xapproxsplr)) +
                     (Matrix::t(S)%*%u)%*%(t(u)%*%xapproxsplr)))

```


Both `svd` and `power_method_dense` result in errors, as the matrix is too large and will cause storage issues. Thus `power_method_sparse_plus_low_rank` is what has to be used to compute the result.

